---
title: "CMTH 642 - Assignment 2"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# USDA Clean Data

We uplodaded the clean csv file generated from Assignment 1 (USDA_Clean.csv).
Please download and load it to your workspace.

```{r}
library(dplyr)
USDAclean = read.csv("C:\\Users\\Vivian\\Desktop\\USDA_Clean.csv")
attach(USDAclean) ## Optional
# attch() function helps you to access USDA_Clean without the need of menioning it.
# For example, you can use Calories instead of USDA_Clean$Calories
View(USDAclean)
str(USDAclean)
```


# Visualization of Feature Relationships

We have used a function panel.cor() inside pair() to show the correlations among different features. The only line you should complete is the line that you assign a value to **USDA_Selected_Featuers**. Research how can you select multiple columns from a dataframe to use it inside pair() function. 

A) Show the relationship among *Calories*, *Carbohydrate*, *Protein*, *Total Fat* and *Sodium*. **(5 p)**

B) Describe the correlations among **Calories**  and other features. **(5 p)**

Hint: We usually interpret the absolute value of correlation as follows:

.00-.19 *very weak*

.20-.39 *weak*

.40-.59 *moderate*

.60-.79 *strong*

.80-1.0 *very strong*


```{r}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor) 
{
    usr <- par("usr"); on.exit(par(usr)) 
    par(usr = c(0, 1, 0, 1)) 
    r <- abs(cor(x, y)) 
    txt <- format(c(r, 0.123456789), digits=digits)[1] 
    txt <- paste(prefix, txt, sep="") 
    if(missing(cex.cor)) cex <- 0.8/strwidth(txt) 
 
    test <- cor.test(x,y) 
    # borrowed from printCoefmat
    Signif <- symnum(test$p.value, corr = FALSE, na = FALSE, 
                  cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1),
                  symbols = c("***", "**", "*", ".", " ")) 
 
    text(0.5, 0.5, txt, cex = cex * r) 
    text(.8, .8, Signif, cex=cex, col=2) 
}
# Assign a value USDA_Selected_Featuers that represents
# "Calories","Carbohydrate","Protein","TotalFat", "Sodium" columns
###################################################
##### Complete code here and uncomment it
USDA_Selected_Featuers = data.frame(select(USDAclean, Calories, Carbohydrate, Protein, TotalFat, Sodium))
###################################################

#### Uncomment the following line when you assign USDA_Selected_Featuers to show the results
pairs(USDA_Selected_Featuers, lower.panel=panel.smooth, upper.panel=panel.cor)

# There is a positive correlation between between Calories and TotalFat. The scatterplot shows the data points tightly packed together and increasing positively, which is also reinforced by the value of correlation at 0.8. A value of 0.8 would suggest that it is a very strong positive correlation and a signaficance level of 3 stars suggests it is highly significant.
# The next highest correlation value would be Calories vs Carbohydrates at 0.42 which is also highly significant at 3 stars. This would be considered a moderate level correlation between the two variables.
# A third relationship that should be looked at would be Carbohydrate vs Protein. It is siginificant at 3 stars with a correlation value of 0.3. It is weakly correlated between each other but has a positive correlation. 
# This would make sense because generally fat contains more calories than carbohydrates or protein and it is more prevalent in the foods we eat. So naturally the more food aka calories we intake, the more fat we intake as well. The plot of Carbs vs Calories is like a mountain, this would make sense as well because foods that are considered carbohydrates can vary in calories. Calories can take the form of different types such as carbs, protein, and fat. An item like avocados are low in carbs but high in calories because it contains alot of beneficial fats. There are also alot of high carb high calorie foods because of the amount of overprocessed foods available. Lastly carbohydrates vs proteins is for the most part shows little correlation, judging from the plot most of the data points are flat on the bottom horizontally which we have concluded early because of the 0.3 correlation value. This makes sense because generally foods that have carbs do not have protein and foods that contain protein dont have carbs. When protein is low, thats were the majority of data points are, just varying in carbohydrate content. When protein is high, the carbohydrate content is low.
# Another thing to note is that sodium has almost no correlation to any of the other variables, it is a very weak correlation. 


```

## Regression Model on USDA Clean Data
Create a Linear Regression Model (lm), using **Calories** as the dependent variable, and *Carbohydrate*, *Protein*, *Total Fat* and *Sodium* as independent variables. **(10 p)**
```{r}
CalorieRegression = lm(Calories ~ Carbohydrate + Protein + TotalFat + Sodium, data = USDAclean)
summary(CalorieRegression)

```

## Analyzing Regression Model
A) In the above example, which independent feature is less significant? (Hint: Use ANOVA) **(5 p)**


```{r}
anova(CalorieRegression)

```
B) Which independent variable has the strongest positive predictive power in the model? (Hint: Look at the coefitients calculated for each independant variable) **(5 p)**
```{r}
# The independent feature that is less significant is Sodium. The value with the strongest positive predictive power is TotalFat at 8.78.

```
## Calories Prediction
A new product is just produced with the following data:

"Protein"	"TotalFat"	"Carbohydrate"	"Sodium"	"Cholesterol"

0.1       40        425             430       75 

"Sugar"	"Calcium"	"Iron"	"Potassium"	"VitaminC"	"VitaminE"	"VitaminD"

NA      42        NA      35			    10			    0.0			    NA

A) Based on the model you created, what is the predicted value for **Calories** ? **(5 p)**

B) If the *Sodium* amount increases 101 times from 430 to 43430 (10000% increase), how much change will occur on Calories in percent? Can you explain why? **(5 p)**

```{r}
predict(CalorieRegression, data.frame(Protein = 0.1, TotalFat = 40, Carbohydrate = 425, Sodium = 430))
predict(CalorieRegression, data.frame(Protein = 0.1, TotalFat = 40, Carbohydrate = 425, Sodium = 43430))
# The predicted value for Calories would be 1943.
# The percentage change in calories if we increased Sodium to 43430 is 0.72%. This is equivalent to the increase in sodium(43000) multiplied by the coefficient for sodium which is 0.0003249. This would give us the increase of 13.97 calories which is the difference between 1957.622 - 1943.65.

```

# Wilcoxon Tests

## Research Question: Does illustrations improve memorization? 
A study of primary education asked elementaty school students to retell two book articles that they read earlier in the week. The first (Article 1) had no picutres, and the second (Article 2) illustrated with pictures. An expert listened
to recordings of the students retelling each article and assigned a score for
certain uses of language. Higher scores are better. Here are the data for five readers in a this study:

Student     1     2     3     4     5

Article 1   0.40  0.72  0.00  0.36  0.55

Article 2   0.77  0.49  0.66  0.28  0.38

We wonder if illustrations improve how the students retell an article. 

### What is $H_0$ and $H_a$ ? 

**(10 p)**

```{r}
# $H_0$ : The mean score for illustrations and no illustrations is the same in the population.
# $H_a$ : The mean score for illustrations and no illustrations is not the same in the population.

```

### Paired or Independent design?

Based on your answer, which Wilcoxon test should you use? **(5 p)**
```{r}
# Since the study uses the same students for the first article and the second article, the samples are paired. Therfore, the Wilcoxon Signed rank test should be used. 

```

### Will you accept or reject your Null Hypothesis? ($\alpha = 0.05$)
Do illustrations improve how the students retell an article or not? **(5 p)**

```{r}
Article1 = c(0.4, 0.72, 0.0, 0.36, 0.55)
Article2 = c(0.77, 0.49, 0.66, 0.28, 0.38)

IllustrationTest = wilcox.test(Article1, Article2, paired = TRUE, alternative = "two.sided")
IllustrationTest
# The P-value of the test is 0.8125 which is higher than the than the significance level alpha = 0.05. The difference between the means is not significantly different, therfore we fail to reject the null hypothesis. Based on the data we have, illustrations do not improve how the students retell an article at this time. 
```

## Packaging Problem

Two companies selling toothpastes with the lable of 100 grams per tube on the package. We randomly bought eight toothpastes from each company A and B from random stores. Afterwards, we scaled them using high precision scale. Our measurements are recorded as follows:

Company A: 97.1   101.3   107.8   101.9   97.4    104.5   99.5    95.1

Company B: 103.5  105.3   106.5   107.9   102.1   105.6   109.8   97.2

### Distribution Analysis
Are the distributions of package weights similar for these companies? Are they normally distributed or skewed? **(10 p)** (Hint: Use boxplot)

```{r}
CompanyA = c(97.1, 101.3, 107.8, 101.9, 97.4, 104.5, 99.5, 95.1)
CompanyB = c(103.5, 105.3, 106.5, 107.9, 102.1, 105.6, 109.8, 97.2)

boxplot(CompanyA, CompanyB)
plot(density(CompanyA))
plot(density(CompanyB))
# The distributions are different for these two companies. Company A is normally distributed and Company B has a bimodal distribution skewed to the left. 
```

### Are packaging process similar or different based on weight measurements? 

Can we be at least 95% confident that there is no difference between packaging of these two companies? **(5 p)** 

Can we be at least 99% confident? **(5 p)** 

Please explain.

```{r}
t.test(CompanyA, CompanyB)
# We can not be at least 95% confident because at the 5% significance level, the p-value is not statisitically significant. Neither is the 1% significance level. P-value > alpha levels 0.05 and 0.01.


```
 
# Correlation
Plot and see the relationship between "cylinder" (cyl) and "weight" (wt) of the cars from mtcars dataset. 
A) Can you see any patterns of correlation between these two variable? **(5 p)**

```{r}
attach(mtcars)
plot(cyl, wt)
# Based on the plot, the correlation between these two variables is that the more cylinders a car has, the heavier it will be. 

```

B) What is the best description for "cyl" and "wt" variables? (Ratio, Ordinal, Interval, or Categorical) **(5 p)**

```{r}
# Weight would be an interval variable. This is because weight carries an order and the differences between values matter, weight can also be zero. Number of cylinders(cyl) is ordinal as the order matters but the differences between each are not clearly defined. Six cylinders is better than four cylinders generally but it is unclear just how much better the extra two cylinders are. 

```

C) Based on the description of the "cyl" and "wt" variables, should you use "Pearson" or "Spearman" correlation? Find the correlation between these two variables. **(10 p)**

```{r}
# Because cyl is an ordinal variable, it would be better to use Spearman correlation between these two variables.
# The spearman correlation between these two variables is 0.8577.
cor(cyl, wt, method = "spearman")


```
