---
title: 'CMTH 642 Data Analytics: Advanced Methods Assignment 1 - Derick Tung'
output:
  html_document:
    df_print: paged
---

### 1. Read the csv files in the folder. (4 points)
```{r}
USDA_Macro = read.csv("C:\\Users\\Vivian\\Downloads\\USDA_Macronutrients.csv")
USDA_Micro = read.csv("C:\\Users\\Vivian\\Downloads\\USDA_Micronutrients.csv")
head(USDA_Macro)
head(USDA_Micro)
```

### 2. Merge the data frames using the variable "ID". Name the Merged Data Frame "USDA". (4 points)
```{r}
USDA = merge(USDA_Macro, USDA_Micro, by = "ID")
head(USDA)
```

### 3. Check the datatypes of the attributes. Delete the commas in the Sodium and Potasium records. Assign Sodium and Potasium as numeric data types. (6 points)
```{r}
sapply(USDA, class)

USDA$Potassium = gsub(",", "", USDA$Potassium)
USDA$Sodium = gsub(",", "", USDA$Sodium)
USDA$Potassium = as.numeric(USDA$Potassium)
USDA$Sodium = as.numeric(USDA$Sodium)

sapply(USDA, class)

```

### 4. Remove records (rows) with missing values in more than 4 attributes (columns). How many records remain in the data frame? (6 points)
```{r}
USDA2 = USDA[rowSums(is.na(USDA)) < 5, ]

USDA2
"6887 records remain in the data frame."
```

### 5. For records with missing values for Sugar, Vitamin E and Vitamin D, replace missing values with mean value for the respective variable. (6 points)
```{r}
sugarmean = mean(USDA2$Sugar, na.rm = TRUE)
USDA2$Sugar[is.na(USDA2$Sugar)] = sugarmean

vitemean = mean(USDA2$VitaminE, na.rm = TRUE)
USDA2$VitaminE[is.na(USDA2$VitaminE)] = vitemean

vitdmean = mean(USDA2$VitaminD, na.rm = TRUE)
USDA2$VitaminD[is.na(USDA2$VitaminD)] = vitdmean

USDA2
```

### 6. With a single line of code, remove all remaining records with missing values. Name the new Data Frame "USDAclean". How many records remain in the data frame? (6 points)
```{r}
USDAclean = USDA2[complete.cases(USDA2), ]
"6310 records remain in USDAclean."
USDAclean
```

### 7. Which food has the highest sodium level? (6 points)
```{r}
USDAclean$Description[which.max(USDAclean$Sodium)]
"Table Salt has the highest sodium level"
```

### 8. Create a histogram of Vitamin C distribution in foods, with a limit of 0 to 100 on the x-axis and breaks of 100. (6 points)
```{r}
hist(USDAclean$VitaminC, main = "Vitamin C distribution in foods", xlim = c(0,100), breaks = 100, xlab = "Vitamin C")
```

### 9. Create a boxplot to illustrate the distribution of values for TotalFat, Protein and Carbohydrate. (6 points)
```{r}
boxplot(USDAclean$TotalFat, main = "Total Fat Distribution", ylab = "Total Fat")

boxplot(USDAclean$Protein, main = "Protein Distribution", ylab = "Protein")

boxplot(USDAclean$Carbohydrate, main = "Carbohydrate Distribution", ylab = "Carbohydrate")

```

### 10. Create a scatterplot to illustrate the relationship between a food's TotalFat content and its calorie content. (6 points)
```{r}
plot(USDAclean$TotalFat, USDAclean$Calories, main = "Relationship between Total fat content and calorie content", ylab = "Calories", xlab = "Total Fat")
```

### 11. Add a variable to the data frame that takes value 1 if the food has higher sodium than average, 0 otherwise.Call this variable HighSodium. Do the same for High Calories, High Protein, High Sugar, and High Fat. How many foods have both high sodium and high fat? (8 points)
```{r}
USDAclean["HighSodium"] = ifelse(mean(USDAclean$Sodium) < USDAclean$Sodium, 1, 0)
USDAclean["HighCalories"] = ifelse(mean(USDAclean$Calories) < USDAclean$Calories, 1, 0)
USDAclean["HighProtein"] = ifelse(mean(USDAclean$Protein) < USDAclean$Protein, 1, 0)
USDAclean["HighSugar"] = ifelse(mean(USDAclean$Sugar) < USDAclean$Sugar, 1, 0)
USDAclean["HighFat"] = ifelse(mean(USDAclean$TotalFat) < USDAclean$TotalFat, 1, 0)

USDAclean
table(USDAclean$HighSodium, USDAclean$HighFat)
"The number of foods with high sodium and high fat is 644."
```

### 12. Calculate the average amount of iron, sorted by high and low protein. (8 points)
```{r}
tapply(USDAclean$Iron, USDAclean$HighProtein, mean)
```

### 13. Create a script for a "HealthCheck" program to detect unhealthy foods. Use the algorithm flowchart below as a basis for this script. (8 points)
```{r}
require(jpeg)
img<-readJPEG("HealthCheck.jpg")
plot(1:4, ty = 'n', ann = F, xaxt = 'n', yaxt = 'n')
rasterImage(img,1,1,4,4)
```
```{r}
HealthCheck = function(x,y,z){
  ifelse(x==1,
         ifelse(y==1,
                ifelse(z==1, "Fail", "Pass"), "Pass"), "Pass")
}
```

### 14. Add a new variable called HealthCheck to the data frame using the output of the function. (8 points)
```{r}
USDAclean["HealthCheck"] = HealthCheck(USDAclean$HighSodium, USDAclean$HighSugar, USDAclean$HighFat)

```

### 15. How many foods in the USDAclean data frame fail the HealthCheck? (8 points)
```{r}
x = sum(USDAclean$HealthCheck == "Fail")
x
```

### 16. Save your final data frame as "USDAclean_ [your last name]" (4 points)
```{r}
USDAclean_Tung = USDAclean
USDAclean_Tung
```


